---
title: "506 final project"
author: "xue luo"
date: "2/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Purpose of the project
### This project is to study the chronological change of human warefare around the world and see if the development vary by different continents. I will first study the global trend of popoluation size, life span, and income increase, and then make comparison among continents.

# view and summarize data
### This global dataset contains demographic information of 197 countries (including 1 as other) with life expasion, population, income and region from 1800 to 2015. In total there are 41284 observations, and each country has 216 observations, and the "other" country has 19900.

```{r}
gapminderdata<- read.csv("/Users/luoxue/Desktop/HU/ANLY 506 Exploratory Data Analytics/final_project/Data/gapminder.csv")
View(gapminderdata)
str(gapminderdata)
summary(gapminderdata)
summary(gapminderdata$Country)
length(unique(gapminderdata$Country))
```

